---
title: "Course Project 1"
author: "TIm Annable"
date: "November 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###About
This project is a component of the Reproducable Research course in the Data Science Specialization.  It will demonstrate how to load and process data in a literate statistical programming fashion.

###Loading and reading data
Data is downloaded directly from the link in the assignment.  It is 17,568 obeservations

Variables include steps, date, and interval.

* **Steps**: Number of steps in a 5 minute interval

* **Date**: The date on which the measurement is taken

* **Interval**:  Identifier for the 5 minute interval in which the measurement is taken

```{r}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
unzip(temp)
unlink(temp)
activitydata <- read.csv("activity.csv")
```
###Create Histogram for total steps per day
This will group stepb sy day, produce a histogram, and calculate summary statistics

```{r}
daily_steps <- aggregate(steps ~ date, activitydata, sum)
hist(daily_steps$steps, main = paste("Steps per day"),xlab="Number of Steps")


```
### Display the mean and median number of steps taken each day
```{r}
themean <- mean(daily_steps$steps)
themedian <- median(daily_steps$steps)
```



The `mean` is `r themean` and the `median` is `r themedian`

###Create time series plot of the average number of steps taken

```{r}
interval_steps <- aggregate(steps ~ interval, activitydata, mean)
plot(interval_steps$interval,interval_steps$steps,type="l",xlab="Interval", ylab="Steps", main="Average steps taken per interval")


```

###The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max_inter <- interval_steps[which.max(interval_steps$steps),1]

```

The interval that, on average, contains the larges number of chunks is `r max_inter`.

###Code to describe and show a strategy for imputing missing data

Imputation of the average for that particular interval.  Whatever the average for that interval was for the non-missing vlaues will be imputed. Drop 2012-10-01 becasue the data would calculate average incorrectly

```{r}
missing_data <- sum(!complete.cases(activitydata))
imputed_data <- transform(activitydata, steps = ifelse(is.na(activitydata$steps),interval_steps$steps[match(activitydata$interval,  interval_steps$interval)],activitydata$steps))

imputed_data[as.character(imputed_data$date) == "2012-10-01",1] <-0

```

###HIstogram of the total number of steps taken each day after imputation

```{r}
daily_steps_2 <- aggregate(steps ~ date, imputed_data, sum)
hist(daily_steps_2$steps, main = paste("Steps per day"),xlab="Number of Steps")

```

###Create a panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays),"Weekday", "Weekend"))
interval_steps_imputed <- aggregate(steps ~ interval + dow, imputed_data, mean)
library(lattice)
xyplot(interval_steps_imputed$steps ~ interval_steps_imputed$interval | interval_steps_imputed$dow,  main="Avg Steps per day by interval", type="l", xlab="Interval", ylab="Steps", layout=(c(1,2)))


```
